Service-Oriented ARC Storage; \\An Overview and Design

### 2008-12-05 doc/storage_system/ARCStorage_src/ARCStorage.tex
proof-reading

\begin{document}
\pagestyle{headings} 
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$ 
\\$^{*}$Dept. of Information Technology, Div. of Scientific Computing, Uppsala
University,\\ Box~337, SE-75105~Uppsala, Sweden
 \\ $^{\dagger}$University of Oslo, Center for Information Technology,\\
P.~O.~Box~1059, Blindern, N-0316~Oslo, Norway
  \\ $^{\ddagger}$University of Oslo, Dept. of Physics, P.~O.~Box~1048,
Blindern, N-0316~Oslo, Norway
  \\  $^{\mathsection}$Dept. of Physics and Astronomy, Div. of Nuclear and
Particle Physics,\\ Uppsala University, Box~535, SE-75121~Uppsala, Sweden \\
$^{\mathparagraph}$Institute of National Information and Infrastructure
Development\\ NIIF/HUNGARNET, Victor~Hugo~18-22, H-1132~Budapest, Hungary\\ \\ 
j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

\maketitle             
%% \begin{abstract}
%% The KnowARC project will deliver the next generation Advanced Resource
%% Connector Fall  2009. A major component of this
%% service oriented Grid middleware will be its storage system, the
%% ARC Storage, of which a
%% prototype is due Spring 2009. We will in this paper present an
%% overview of the storage system, its design and architecture. We will
%% also present some first, proof-of-concept test results, deploying the
%% the storage with a geographically distribution among three different countries.
%% ~\\~\\~
%% \end{abstract}
\begin{abstract}
There is an ever increasing need to utilize geographically distributed
hardware resources, both in terms of CPU and in terms of storage. The
service oriented architecture provides a natural framework for
managing these resources. The next generation Advanced Resource
Connector (ARC) is a service oriented Grid solution that will provide
the middleware to represent distributed resources in one simple
framework. In this paper, we will present an overview of the ARC
storage system, itself a set of services providing a self-healing,
flexible Grid storage solution. We will also present some first
proof-of-concept test results, with a deployment of the storage system
distributed across three different countries. 
~\\~\\~
\end{abstract}
%-----------------------------------------------------------
\section{Introduction}
\label{Introduction}

The challenge of building a reliable, self-healing, fault-tolerant,
consistent data management system in a web scale is an interesting
task. Making the system work in a heterogeneous, distributed
environment like the Grid is even more interesting. An increasing
number of applications demand not only increased CPU power, but also
vast amounts of storage space. The required storage space is not only
restricted to the duration in which the application runs; the data
should often be available for years afterwards, in certain cases even
for decades. Nowadays, we can easily find single Grid jobs which produce
gigabytes or even terabytes of data, ramping up the requirements of
storage systems to the petabyte scale and beyond. Hard drive capacity is
still increasing and storage raids can handle failing hard drives through 
data replication. However, even a designated storage site cannot guarantee 
24/7 availability or long term security against catastrophic losses of data. 
The need for a distributed, self-healing storage system is evident. To
make the storage system useable to Grid users, the system must provide
reliable and secure file transfer protocols, a cataloging system and
secure storage of the data. Several projects and designs have emerged
to address such challenges~\cite{Hoschek00datamanagement,DengWang}.

In the advent of the next generation of the Advanced Resource Connector (ARC)
Grid middleware~\cite{arc} (new release due during fall 2009), we present 
the ARC storage system~\cite{ARCStoragedesigndoc}. This distributed
storage system is designed to provide an easy
to use, flexible and scalable system that can offer native storage and
at the same time provide access to third-party solutions like
dCache~\cite{dCache,DSSWithdCache}  by
using the same, uniform interface. Being part of ARC, the
storage system is based on a service
oriented architecture, in which each major component of the
system runs as a separate service within the ARC Hosting Environment
Daemon (HED)~\cite{HEDdesigndoc}. The HED service container gives the
capability of flexible replacement of the components as well as the
possibility to introduce modifications in the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

This paper is organized as follows. After describing
related work in Section \ref{Related Work}, we give a bird's eye
view of the next generation Advanced Resource Connector (ARC) in
Section \ref{The Advance Resource Connector}. An overview of the
ARC Storage is given in Section \ref{The ARC Storage System}, while
the architecture of the storage system is elaborated in Section
\ref{Architecture of the ARC Storage}. In Section \ref{Testing and
  Discussion} we give some early, proof-of-concept results, before 
concluding in Section \ref{Conclusion and Future Work}.

@@

the petabyte scale \cite{Bigtable}. It is currently used within
Google in projects like web indexing, Google Earth and Google

@@

and olbd components offers a cluster storage designed for low latency

@@

for high latency more suitable for the Grid environment.

@@

To deliver the non-trivial quality of services required by the Grid,
there are a number of services running inside the HED. For example, Grid
job execution and management is handled by the A-REX service 
\cite{AREXdesigndoc}, policy
decisions are taken by the Charon service, the ISIS service is
responsible for the information indexing, batch job submission is
handled by the Sched service, etc. In the later sections our discussion
will focus on the architecture and the design of another major set
of services, i.e., the ARC storage.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.8\columnwidth]{arc1-storage-services-with-gateway.pdf}
  \end{center}
  \caption{Schematic of the ARC Storage architecture. The figure shows
    the main services of ARC Storage and the communication channels. The B stands for Bartender, the S Shepherd, the L Librarian, the G Gateway and the A-H stands for A-Hash. The straight lines denotes the communication between services.}
  \label{fig:arc1_storage_service}
\end{figure}

Being based on a service oriented architecture, the ARC
Storage consists of a set of services as shown in Fig.
\ref{fig:arc1_storage_service}. The services are as follows: The Bartender (B)
provides the high-level interface to the user; the Librarian (L)
handles the entire storage namespace, using the A-Hash (A-H) as a metadatabase; the
Gateway (G) provides access to third-party storage systems; and
the Shepherd (S) is the frontend for the physical storage node. 
See Section \ref{Architecture of the ARC Storage} for a detailed
discussion of the different services.   
The services communicate with each other through the
Message Chain Components in HED. The communication channels are
depicted by straight lines in Fig. \ref{fig:arc1_storage_service}.

@@

In the default implementation, the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC storage system in the foreseen
future is restricted to provide support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

@@

External grid middleware components can
access the ARC storage system by using ARC data service
interfaces directly. ARC will also provide interface
components that communicate via standard protocols like SRM, which
will provide a single access point to the whole system.

@@

In a service oriented architecture
the role of each service is well defined. Available
objects\footnote{Files, collections, mount points, etc.} in the
system are identified by unique global IDs. These IDs are
categorized according to the object type:

@@

Each object in the ARC Storage has a globally
unique ID. A collection contains files and other collections,
and each of these entries has a name unique within the collection very
much like entries in a standard directory on a local filesystem. Besides
files and collections the ARC Storage has a third type of entry
called mount-points, which are references to the third-party storages
within the global namespace.

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas have assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has wrong checksum, or the Shepherd claims it has no
such replica) and `\textbf{creating}' (if the replica is in
the state of uploading).

In the following the details of the
architecture are presented in three parts: First we will discuss the
details of the core components, second we will discuss some of the
important features provided by the system, and third we will discuss the security model.

@@

\subsubsection{Replication}
\label{Replication}
Shepherds periodically ask the Librarian if the file of the replica stored on
its storage node has enough replicas. If the file does not have
enough replicas, the Shepherd informs the Bartender and the Bartender initiates
a put request that returns a Transfer URL to the Shepherd. The
Shepherd finally uploads the new replica. The Shepherds which gets the new replica notifies the
Librarian that the replica is alive. The Librarian then add this to
the corresponding file.

@@

Due to the unpredictable nature of the Grid environment, it is essential to
have some degree of  automatic recovery system in case of unexpected
failures. Fault tolerant behavior is required both at the level of
metadata and on the level of physical storage. While the work on fault-tolerant metadata is still in
progress, the following two
scenarios will explain the currently available recovery mechanism for
physical storage: 
\begin{itemize}
\item In the case of a file having invalid checksum, the Shepherd
immediately informs the Librarian and the Librarian changes the state of the given replica to
`\textbf{invalid}'. To recover its replica, the Shepherd contacts a
Bartender and asks for another replica of the file. The Bartender chooses a valid
replica, initiates a file transfer from a Shepherd having the replica,
and returns the TURL to the Shepherd with the invalid replica. When the
Shepherd has recieved the replica and compared the checksum, it notifies the
Librarian that the replica is alive again.  
\item In the case of a Shepherd going offline, a Librarian will,
  as mentioned earlier, notice the lack of heartbeats and invalidate all the
  replicas, initiating new replication for all the files stored in
  this storage node. However, if the Shepherd
  again comes online, there will evidently be more replicas than
  needed. The first Shepherd to notice this will set its replica's
  state to `\textbf{thirdwheel}', i.e., obsolete. At the next occasion,
  the Shepherd will remove the replica, if and only if it has the only
  `\textbf{thirdwheel}' replica of this file. If there are more
  replicas with this state, all replicas will be set back to
  `\textbf{alive}' and the process is repeated. This scenario will be discussed further in Section
  \ref{Testing and Discussion}.
\end{itemize}
\subsubsection{Client tools }
\label{Client extension using FUSE module}

Being the only part a user will (and should) see from a storage
system, the client tools are an important part of the ARC
Storage. Currently ARC supports two ways of accessing the
storage solution:
\begin{itemize}
\item The \textbf{Command-line Interface} (CLI) provides
  access to the storage through the methods \texttt{stat}, \texttt{makeCollection},
  \texttt{unmakeCollection}, \texttt{putFile}, \texttt{getFile}, \texttt{delFile},
  \texttt{list} and \texttt{move}. Methods for modifying
  access and ownership will be available in the near future. The CLI
  assumes a relatively high level of computer competence
  from the user. However, the CLI, being a stand-alone tool, can be
  used to access ARC Storage from any computer (also including
  Microsoft Windows PCs) that has network access and
  a Python installation.
\item The \textbf{FUSE module} provides a high-level access to the
  ARC Storage. Filesystem in Userspace (FUSE) \cite{FUSE} provides a simple
  library and a kernel-userspace interface. Using FUSE and the ARC
  Storage Python interface, the FUSE module allows users to mount the
  ARC Storage namespace into the local namespace. This way, the user
  can use her/his favorite file manager to access her/his files and
  collections. The FUSE module provides most of the features provided
  by the CLI, with the exception of modifying some non-posix metadata.
\end{itemize}
Its worth mentioning that the client tools access the storage system
through the Bartender only. Currently upload and download is realized
through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

@@

Gateways are used to communicate with the external storage
managers. While designing this service, care was taken to: 
\begin{itemize}
\item Retain the transparency of the global namespace while using the external storage systems.
\item  Develop a protocol oriented  service i.e., all the external
  storage managers which support a certain protocol should be handled
  using the corresponding gateway service.  
\end{itemize}
This approach provides flexibility while avoiding multiple Gateway
services for different storage managers. Currently, the available
Gateway service is based on the gridftp protocol.

When the user request is made, the Librarian provides the metadata related
to the request to the Bartender. Requests can be related to
files, collections or external mount-points. In case of creating
external mount-points, the Bartender contacts the Librarian to store
the mount-point for later use. In the case where a request is related
to the downloading of files from the external store, the Gateway service first checks the status of the file
and then sends the Transfer URL (TURL) to the client via the
Bartender. Using this TURL, the client can directly get the file from
external store. %% For the experiments we use dCache with gridftp
%% protocol as an external storage system.

@@

As is the case for all openly accessible web services, the security
model is of crucial importance for the ARC Storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization.
\begin{itemize}
\item The \textbf{inter-service authorization} maintains the integrity of the
  internal communication between services. There is a lot of
  communication between the services of the ARC Storage. The
  Bartenders send requests to the Librarians and the Shepherds, the
  Shepherds communicate with the Librarians and the Librarians talk with
  the A-Hash. If any of these services is compromised or a new
  rogue service gets inserted in the system, the security of the
  entire system is compromised. To enable trust between the services,
  they need to know each other's Distinguished Names (DNs). This way,
  a rogue service would need to obtain a certificate with that exact
  DN from some trusted Certificate Authority (CA).
\item The \textbf{transfer-level authorization} handles the
  authorization in the cases of uploading and downloading files. When
  a transfer is requested, the Shepherd will provide a one-time
  Transfer URL (TURL) which the client can connect to. In the
  current architecture, this TURL is world-accessible. This may not seem
  very secure at first. However, provided that the TURL has a very long,
  unguessable name, that it is transfered to the user in a secure way
  and that it can only be accessed once before it is deleted, the
  chance of being compromized is very low.
\item The \textbf{high-level authorization} considers the access
  policies for the files and collections in the system. These policies
  are stored in A-Hash, in the metadata of the corresponding file or
  collection, providing a fine-grained security in the system.
\end{itemize}
The communication with and within the storage system is realized
through HTTPS with standard X.509 authentication.

@@

In our test deployment, we used two nodes from Uppsala
Multidisciplinary Center for Advanced Computational Science (UPPMAX), Sweden, 
%each with a 3.00GHz Dual Core Intel(R) Pentium(R) 4 CPU and 4 GB RAM, 
three nodes from the Center for Information Technology (USIT) at the
University of Oslo, Norway, 
%each with two 3.40GHz Dual core Intel(R) Xeon(TM) CPUs with 2GB RAM
and one node from National Information Infrastructure Development Institute (NIIF)
Hungary. %with a 2.00GHz Dual Core Intel(R) Pentium(R) 4 CPU with 4 GB RAM.

@@

\begin{itemize}
\item Test 1: The distribution of data after uploading 10 large size (1 GB)
  files with one replica each.  
\item Test 2: The distribution of files after uploading 100 small size (1 kB)
  files with one replica each, and with simultanious uploading
\item Test 3: The distribution of data after repeating Tests 1 and 2 with three replicas.  
\item Test 4: System behavior while some of the shepherds are offline.   
\end{itemize}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\columnwidth]{FileDistribution}
\caption{The distribution of replicas over the geographically distributed
  storage nodes. Green bars shows the distribution in the case where
  we upload 100 files requesting 3 replicas where as orange bars show
  the distribution after uploading 100 files without replication. }
\label{fig:FileDistribution}
\end{figure}

@@

Table \ref{table:load distribution} shows the disk usage after Tests
1, 2 and 3. It is worth noticing here that the bandwidth to NIIF was
significantly lower than between UPPMAX and USIT. If the bandwith is
saturated and the storage node is busy, e.g., checksumming a
large file, the heartbeat from this Shepherd may be delayed, causing the
Bartender not to choose this Shepherd for the next replica. This may explain the
relatively low storage load on the  NIIF storage node.

@@

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking one Shepherd out of the system, and later reinserting it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished; and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted and that the storage load evens out 
again when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. We also see that the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaniously
starts uploading replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

@@

The proposed system is still in an early phase of development, but our test
results demonstrate that the architecture is robust enough to handle
the challenges for distributed large scale storage. Much
effort is required to make the system production
ready. However, we believe that continuing in the same direction
will enable us to provide a persistent  and flexible storage system
which can fulfill the needs of even the most demanding scientific community.

>>>>

\begin{document}
\pagestyle{headings} 
\title{Service-Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$ 
\\$^{*}$Dept. of Information Technology, Div. of Scientific Computing, Uppsala
University,\\ Box~337, SE-75105~Uppsala, Sweden
 \\ $^{\dagger}$University of Oslo, Center for Information Technology,\\
P.~O.~Box~1059, Blindern, N-0316~Oslo, Norway
  \\ $^{\ddagger}$University of Oslo, Dept. of Physics, P.~O.~Box~1048,
Blindern, N-0316~Oslo, Norway
  \\  $^{\mathsection}$Dept. of Physics and Astronomy, Div. of Nuclear and
Particle Physics,\\ Uppsala University, Box~535, SE-75121~Uppsala, Sweden \\
$^{\mathparagraph}$Institute of National Information and Infrastructure
Development\\ NIIF/HUNGARNET, Victor~Hugo~18-22, H-1132~Budapest, Hungary\\ \\ 
j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

\maketitle             
%% \begin{abstract}
%% The KnowARC project will deliver the next generation Advanced Resource
%% Connector Fall  2009. A major component of this
%% service-oriented Grid middleware will be its storage system, the
%% ARC Storage, of which a
%% prototype is due Spring 2009. We will in this paper present an
%% overview of the storage system, its design and architecture. We will
%% also present some first, proof-of-concept test results, deploying the
%% the storage with a geographically distribution among three different countries.
%% ~\\~\\~
%% \end{abstract}
\begin{abstract}
There is an ever increasing need to utilize geographically distributed
hardware resources, both in terms of CPU and in terms of storage. The
service-oriented architecture provides a natural framework for
managing these resources. The next generation Advanced Resource
Connector (ARC) is a service-oriented Grid solution that will provide
the middleware to represent distributed resources in one simple
framework. In this paper, we will present an overview of the ARC
storage system, itself a set of services providing a self-healing,
flexible Grid storage solution. We will also present some first
proof-of-concept test results from a deployment of the storage system
distributed across three different countries. 
~\\~\\~
\end{abstract}
%-----------------------------------------------------------
\section{Introduction}
\label{Introduction}

The challenge of building a reliable, self-healing, fault-tolerant,
consistent data management system at the web-scale is an interesting
task. Making the system work in a heterogeneous, distributed
environment like the Grid is even more interesting. An increasing
number of applications demand not only increased CPU power, but also
vast amounts of storage space. The required storage space is not only
restricted to the duration in which the application runs; the data
should often be available for years afterwards, in certain cases even
for decades. Nowadays, we can easily find single Grid jobs which produce
gigabytes or even terabytes of data, ramping up the requirements of
storage systems to the petabyte-scale and beyond. Hard drive capacity is
still increasing and storage raids can handle failing hard drives through 
data replication. However, even a designated storage site cannot guarantee 
24/7 availability or long-term security against catastrophic losses of data. 
The need for a distributed, self-healing storage system is evident. To
make the storage system useable to Grid users, the system must provide
reliable and secure file transfer protocols, a cataloging system and
secure storage of the data. Several projects and designs have emerged
to address such challenges~\cite{Hoschek00datamanagement,DengWang}.

In the advent of the next generation of the Advanced Resource
Connector (ARC) Grid middleware~\cite{arc} (new release due during
fall 2009), we present the ARC storage
system~\cite{ARCStoragedesigndoc}. This distributed storage system is
designed to provide an easy to use, flexible and scalable system that
can offer native storage and at the same time provide access to
third-party solutions like dCache~\cite{dCache,DSSWithdCache} by using
the same, uniform interface. Being part of ARC, the storage system is
based on a service-oriented architecture, in which each major
component of the system runs as a separate service within the ARC
Hosting Environment Daemon (HED)~\cite{HEDdesigndoc}. The HED service
container gives the capability of flexible replacement of the
components as well as the possibility to introduce modifications in
the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

This paper is organized as follows. After describing
related work in Section \ref{Related Work}, we give a bird's eye
view of the next-generation Advanced Resource Connector (ARC) in
Section \ref{The Advance Resource Connector}. An overview of the
ARC Storage is given in Section \ref{The ARC Storage System}, while
the architecture of the storage system is elaborated in Section
\ref{Architecture of the ARC Storage}. In Section \ref{Testing and
  Discussion} we give some early, proof-of-concept results, before 
concluding in Section \ref{Conclusion and Future Work}.

@@

the petabyte-scale \cite{Bigtable}. It is currently used within
Google in projects such as web indexing, Google Earth and Google

@@

and olbd components offers a cluster storage designed for low latency,

@@

for high latency and is more suitable for the Grid environment.

@@

To deliver the non-trivial quality of services required by the Grid,
there are a number of services running inside the HED. For example, Grid
job execution and management is handled by the A-REX service 
\cite{AREXdesigndoc}, policy
decisions are taken by the Charon service, the ISIS service is
responsible for information indexing, batch job submission is
handled by the Sched service, etc. In the later sections our discussion
will focus on the architecture and the design of another major set
of services, i.e., the ARC storage.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.8\columnwidth]{arc1-storage-services-with-gateway.pdf}
  \end{center}
  \caption{Schematic of the ARC Storage architecture. The figure shows
    the main services of ARC Storage and the communication channels. The B stands for Bartender, the S Shepherd, the L Librarian, the G Gateway and the A-H stands for A-Hash. The straight lines denote the communication between services.}
  \label{fig:arc1_storage_service}
\end{figure}

Being based on a service-oriented architecture, the ARC
Storage consists of a set of services as shown in Fig.
\ref{fig:arc1_storage_service}. The services are as follows: The Bartender (B)
provides the high-level interface to the user; the Librarian (L)
handles the entire storage namespace, using the A-Hash (A-H) as a metadatabase; the
Gateway (G) provides access to third-party storage systems; and
the Shepherd (S) is the frontend for the physical storage node. 
See Section \ref{Architecture of the ARC Storage} for a detailed
discussion of the different services.   
The services communicate with each other through the
Message Chain Components in HED. The communication channels are
depicted by straight lines in Fig. \ref{fig:arc1_storage_service}.

@@

In the default implementation, the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC storage system in the forseeable
future is restricted to providing support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

@@

External grid middleware components can
access the ARC storage system using ARC data service
interfaces directly. ARC will also provide interface
components that communicate via standard protocols such as SRM, which
will provide a single access point to the whole system.

@@

In a service-oriented architecture
the role of each service is well defined. Available
objects\footnote{Files, collections, mount points, etc.} in the
system are identified by unique global IDs. These IDs are
categorized according to the object type:

@@

Each object in the ARC Storage has a globally
unique ID. A collection contains files and other collections,
and each of these entries has a name unique within the collection very
much like entries in a standard directory on a local filesystem. Besides
files and collections, the ARC Storage has a third type of entry
called mount-points, which are references to the third-party storages
within the global namespace.

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas have assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has a wrong checksum, or the Shepherd claims it has no
such replica) and `\textbf{creating}' (if the replica is in
the state of uploading).

In the following, the details of the
architecture are presented in three parts: First we will discuss the
details of the core components, second we will discuss some of the
important features provided by the system, and third we will discuss the security model.

@@

\subsubsection{Replication}
\label{Replication}
Shepherds periodically ask the Librarian if the file of the replica stored on
its storage node has enough replicas. If the file does not have
enough replicas, the Shepherd informs the Bartender and the Bartender initiates
a put request that returns a Transfer URL to the Shepherd. The
Shepherd finally uploads the new replica. The Shepherd which gets the new replica notifies the
Librarian that the replica is alive. The Librarian then adds this to
the corresponding file.

@@

Due to the unpredictable nature of the Grid environment, it is essential to
have some degree of  automatic recovery system in case of unexpected
failures. Fault tolerant behavior is required both at the level of
metadata and on the level of physical storage. While the work on fault-tolerant metadata is still in
progress, the following two
scenarios will explain the currently available recovery mechanism for
physical storage: 
\begin{itemize}
\item In the case of a file having an invalid checksum, the Shepherd
immediately informs the Librarian and the Librarian changes the state of the given replica to
`\textbf{invalid}'. To recover its replica, the Shepherd contacts a
Bartender and asks for another replica of the file. The Bartender chooses a valid
replica, initiates a file transfer from a Shepherd having the replica,
and returns the TURL to the Shepherd with the invalid replica. When the
Shepherd has recieved the replica and compared the checksum, it notifies the
Librarian that the replica is alive again.  
\item In the case of a Shepherd going offline, a Librarian will,
  as mentioned earlier, notice the lack of heartbeats and invalidate all the
  replicas, initiating new replication for all the files stored in
  this storage node. However, if the Shepherd
  again comes online, there will evidently be more replicas than
  needed. The first Shepherd to notice this will set its replica's
  state to `\textbf{thirdwheel}', i.e. obsolete. At the next occasion,
  the Shepherd will remove the replica, if and only if it has the only
  `\textbf{thirdwheel}' replica of this file. If there are more
  replicas with this state, all replicas will be set back to
  `\textbf{alive}' and the process is repeated. This scenario will be discussed further in Section
  \ref{Testing and Discussion}.
\end{itemize}
\subsubsection{Client tools }
\label{Client extension using FUSE module}

Being the only part a user will (and should) see from a storage
system, the client tools are an important part of the ARC
Storage. Currently ARC supports two ways of accessing the
storage solution:
\begin{itemize}
\item The \textbf{Command-line Interface} (CLI) provides
  access to the storage through the methods \texttt{stat}, \texttt{makeCollection},
  \texttt{unmakeCollection}, \texttt{putFile}, \texttt{getFile}, \texttt{delFile},
  \texttt{list} and \texttt{move}. Methods for modifying
  access and ownership will be available in the near future. The CLI
  assumes a relatively high level of computer competence
  from the user. However, the CLI, being a stand-alone tool, can be
  used to access ARC Storage from any computer (also including
  Microsoft Windows PCs) that has network access and
  a Python installation.
\item The \textbf{FUSE module} provides a high-level access to the
  ARC Storage. Filesystem in Userspace (FUSE) \cite{FUSE} provides a simple
  library and a kernel-userspace interface. Using FUSE and the ARC
  Storage Python interface, the FUSE module allows users to mount the
  ARC Storage namespace into the local namespace. This way, the user
  can use her/his favorite file manager to access her/his files and
  collections. The FUSE module provides most of the features provided
  by the CLI, with the exception of modifying some non-posix metadata.
\end{itemize}
It is worth mentioning that the client tools access the storage system
through the Bartender only. Currently upload and download is realized
through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

@@

Gateways are used to communicate with the external storage
managers. While designing this service, care was taken to: 
\begin{itemize}
\item Retain the transparency of the global namespace while using the external storage systems.
\item  Develop a protocol-oriented  service, i.e. all the external
  storage managers which support a certain protocol should be handled
  using the corresponding gateway service.  
\end{itemize}
This approach provides flexibility while avoiding multiple Gateway
services for different storage managers. Currently, the available
Gateway service is based on the gridftp protocol.

When the user request is made, the Librarian provides the metadata related
to the request to the Bartender. Requests can be related to
files, collections or external mount-points. In case of creating
external mount-points, the Bartender contacts the Librarian to store
the mount-point for later use. In the case where a request is related
to the downloading of files from the external store, the Gateway service first checks the status of the file
and then sends the Transfer URL (TURL) to the client via the
Bartender. Using this TURL, the client can directly get the file from
the external store. %% For the experiments we use dCache with gridftp
%% protocol as an external storage system.

@@

As is the case for all openly accessible web services, the security
model is of crucial importance for the ARC Storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization.
\begin{itemize}
\item The \textbf{inter-service authorization} maintains the integrity of the
  internal communication between services. There is a lot of
  communication between the services of the ARC Storage. The
  Bartenders send requests to the Librarians and the Shepherds, the
  Shepherds communicate with the Librarians and the Librarians talk with
  the A-Hash. If any of these services is compromised or a new
  rogue service gets inserted in the system, the security of the
  entire system is compromised. To enable trust between the services,
  they need to know each other's Distinguished Names (DNs). This way,
  a rogue service would need to obtain a certificate with that exact
  DN from some trusted Certificate Authority (CA).
\item The \textbf{transfer-level authorization} handles the
  authorization in the cases of uploading and downloading files. When
  a transfer is requested, the Shepherd will provide a one-time
  Transfer URL (TURL) to which the client can connect. In the
  current architecture, this TURL is world-accessible. This may not seem
  very secure at first. However, provided that the TURL has a very long,
  unguessable name, that it is transfered to the user in a secure way
  and that it can only be accessed once before it is deleted, the
  chance of being compromized is very low.
\item The \textbf{high-level authorization} considers the access
  policies for the files and collections in the system. These policies
  are stored in A-Hash, in the metadata of the corresponding file or
  collection, providing a fine-grained security in the system.
\end{itemize}
The communication with and within the storage system is realized
through HTTPS with standard X.509 authentication.

@@

In our test deployment, we used two nodes from Uppsala
Multidisciplinary Center for Advanced Computational Science (UPPMAX), Sweden, 
%each with a 3.00GHz Dual Core Intel(R) Pentium(R) 4 CPU and 4 GB RAM, 
three nodes from the Center for Information Technology (USIT) at the
University of Oslo, Norway, 
%each with two 3.40GHz Dual core Intel(R) Xeon(TM) CPUs with 2GB RAM
and one node from National Information Infrastructure Development Institute (NIIF),
Hungary. %with a 2.00GHz Dual Core Intel(R) Pentium(R) 4 CPU with 4 GB RAM.

@@

\begin{itemize}
\item Test 1: The distribution of data after uploading 10 large size (1 GB)
  files with one replica each.  
\item Test 2: The distribution of files after uploading 100 small size (1 kB)
  files with one replica each, and with simultanious uploading.
\item Test 3: The distribution of data after repeating Tests 1 and 2 with three replicas.  
\item Test 4: System behavior while some of the Shepherds are offline.   
\end{itemize}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\columnwidth]{FileDistribution}
\caption{The distribution of replicas over the geographically distributed
  storage nodes. Green bars show the distribution in the case where
  we upload 100 files requesting 3 replicas, whereas orange bars show
  the distribution after uploading 100 files without replication. }
\label{fig:FileDistribution}
\end{figure}

@@

Table \ref{table:load distribution} shows the disk usage after Tests
1, 2 and 3. It is worth noticing here that the bandwidth to NIIF was
significantly lower than between UPPMAX and USIT. If the bandwidth is
saturated and the storage node is busy, e.g., checksumming a
large file, the heartbeat from this Shepherd may be delayed, causing the
Bartender not to choose this Shepherd for the next replica. This may explain the
relatively low storage load on the  NIIF storage node.

@@

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking one Shepherd out of the system, and later reinserting it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished, and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted and that the storage load evens out 
again when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. We also see that the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaneously
start uploading and replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

@@

The proposed system is still in an early phase of development, but our test
results demonstrate that the architecture is robust enough to handle
the challenges for distributed large-scale storage. Much
effort is required to make the system production
ready. However, we believe that continuing in the same direction
will enable us to provide a persistent  and flexible storage system
which can fulfill the needs of even the most demanding scientific community.

Some key areas need special attention and effort to make the proposed
storage system even more stable, reliable and consistent:
\begin{itemize}
\item \textbf{Security:} This is still under development. The design
  is almost ready, but it still needs to be implemented and properly
  tested.
\item \textbf{The A-Hash}: This is currently centralized. To avoid a single
  point of failure in the system, and to improve the system
  performance, the A-Hash needs to be distributed.
\item \textbf{Performance optimization:} To make a storage system
  ready for production, one needs to discover and improve on
  possible bottlenecks within the system. As soon as the A-Hash is
  distributed, other possible bottlenecks, such as load-balancing
  and self-healing mechanisms, will be investigated.
\item \textbf{Protocols:} To ease the interoperability with third-party
  storage solutions and clients, the system needs to support storage
  protocols such as SRM and GridFTP. In addition the system will come
  with its own ARC protocol.
\end{itemize}
